
---
title: "3. Data frames in R"
author: "Christian Devenish"
date: "6 Febuary 2019"
draft: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "markup", eval = T)
```

```{r objects, include = F}

v1 <- c(50,20,3,15,43)
s1 <- c("a", "b", "c", "d", "e") 

```


### The **data frame**  a 'table' in R
*Get the worksheet here* [worksheet 3](../resources/w03_dataframes.r)


We can think of a data frame as a collection of vectors of the same length. For example, let's combine two vectors into a data frame

```{r}
# Think back to this vector:
v1
```

How long is it? i.e. how many elements does it have?
```{r}
length(v1)
```

What type (numeric, character) are the elements?
```{r}
class(v1)
```

Think back to this vector:
```{r}
s1
length(s1)
class(s1)
```

### The **data frame**  a 'table' in R
- a basic object to store data in  
  

```{r}
df1 <- data.frame(Type = s1, Value = v1)
df1
```

**Vectors combined into a data frame must be of same length**
**A data frame can contain columns of different data types**
