
---
title: "4. Importing data into R"
author: "Christian Devenish"
date: "6 Febuary 2019"
draft: false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "markup", eval = T)
sparrows <- read.csv("resources/sparrows.csv")
```


### Getting data from an external source
Download the data file here - save it in your course folder: [sparrows data in excel](../resources/sparrow.xls)  
*Get the worksheet here* [worksheet 4](../resources/w04_importingData.r)

Transform an excel file to a csv with *save as* in the *file* menu. Call it *sparrows.csv*

We use `read.csv()` to import into R.

```{r, eval = F}
sparrows <- read.csv("sparrows.csv")

```

Always check your data after importing:

**`str()`** tells us what kind of object it is, and its structure.

e.g. a Data frame with 5 columns, and 20 rows, and the data type of each column:

```{r}

str(sparrows)

```

**`head()`** shows us the first 6 lines of data:

```{r}

head(sparrows)

```

To access a single column from the data frame, we use the `$` symbol between the data frame name and the column name:

**DATA-FRAME-NAME$COLUMN-NAME**

```{r}
sparrows$Culmen

```



