
---
title: "2. Vectors, variables and other objects"
author: "Christian Devenish"
date: "6 Febuary 2019"
draft: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "markup", eval = T)
```

### Vectors, variables and other objects
*Get the worksheet here* [worksheet 2](../resources/w02_vectors.r)


What is a vector? It's an *ordered* collection of data.
For example, it could be a single column of a table 


```
[1] 3.79 63.00 0.53 6.46 11.46 0.29 0.08 0.74
```
Or a single value

```
[1] 3.141593
```

### Types of vector

- Numeric
 - continuous
 - integer

- Character (string)
 - text

- Logical
TRUE or FALSE

### Vector with a single numeric item


```{r }
a <- 1
b <- 5
x <- 100
# note the assign operator: <-  
# this means assign value on the right hand side to the variable/object name on the left
```

```{r }
a
```


```{r eval = TRUE}
x
```


**Naming objects...**
- Don't use *reserved* words:
`? reserved`
- Best not to use function names
- Don't begin with a number
- Sensitive to upper and lower case
- Don't use spaces
- Don't use symbols such as £ $ % ^ & * ( ) # ? < > / | \ [ ] { }
</small>

### Operations with vectors 1


```{r}
a + b
```


```{r}
(a + b) * x
```



```{r}
a / x
```


```{r}
sum(a, b)
```


```{r}
sqrt(x)
```


```{r}
log10(x)
```


### Character vectors

```{r}
a <- "Hello"  # text ALWAYS in "quotes"
b <- "World"
sp1 <- "Phytotoma raimondii"
```

```{r}
a  # the object name is not in quotes
```


```{r}
sp1
```

```
[1] "Phytotoma raimondii"
```

### Text operations


```{r}
# Concatenate text
paste(a, b)
```


```{r}
# and with a separator
paste(a, sp1, sep=", ")
```

### Vectors with more than one element

```{r}
1:10 # generate a series in steps of 1
```


Using the `c()` function to combine numbers into a vector:
```{r}
# combine 5 numbers and assign to a vector called v1
v1 <- c(50,20,3,15,43)
v2 <- 5:10

v1
v2
```

**Vectors only allow one data type (numeric, character, logical)**

### Operations with vectors


```{r}
v1 * 3  # cf. to excel
```


```{r}
v1 + v2
```


```{r}
mean(v1)
```


```{r}
sum(v1)/length(v1)
```


### Getting a value within a vector with indices []

```{r}
v1
```


```{r}
v1[3]  # get the 3rd element of the vector v1
```

```{r}
v1[2:4] # get elements 2 to 4 of v1
```


```{r}
v1[-1] # get all elements of v1 except for the first
```

### Series of letters

```{r}
s1 <- c("a", "b", "c", "d", "e") # note the c() function
s2 <- letters[1:5]  # letters is a ready-made object inside R
s3 <- LETTERS[1:5]  # see ? Constants

s1
s2
s3
```


### Getting an elements in a character vector


```{r}
s1[1]
```

```{r}
s1[4:5]
```

```{r}
sp1[1] # Is this what you expected?
```


```{r}
s1[-2]
```

